---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# UNSEEN

<!-- badges: start -->
<!-- badges: end -->

The UNprecedented Simulated Extreme ENsemble (UNSEEN, [Thompson et al., 2017](https://www.nature.com/articles/s41467-017-00275-3)) approach is an increasingly popular method that exploits seasonal prediction systems to assess and anticipate climate extremes beyond the observed record. The goal of this package is to provide easy evaluation of the UNSEEN method. This package is developed as part of the ECMWF Summer of Weather Code ([ESoWC 2020](https://github.com/esowc/UNSEEN-open)) using in particular for the ECMWF SEAS5 hindcast data.

## Installation

<!-- You can install the released version of UNSEEN from [CRAN](https://CRAN.R-project.org) with: -->

<!-- ``` r -->
<!-- install.packages("UNSEEN") -->
<!-- ``` -->

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("timokelder/UNSEEN")
```
## Example

This is a basic example which shows you how to plot the UNSEEN timeseries and apply the independence, stability and fidelity tests using an example dataset:

```{r}
data("EOBS_UK","SEAS5_UK")
```


```{r example}
library(UNSEEN)
## basic example code

unseen_timeseries(ensemble = SEAS5_UK, obs = EOBS_UK, ylab = 'UK February precipitation (mm/d)')
independence_test(ensemble = SEAS5_UK)
stability_test(ensemble = SEAS5_UK, lab = 'UK February precipitation (mm/d)')
fidelity_test(obs = EOBS_UK$rr, ensemble = SEAS5_UK$tprate)
```

